```{r}
library(ggplot2)
theme_set(theme_bw())
```


# Results

How well do uncertainty models characterize empirical errrors from simulated SWOT data? The answer depends on the variable measured, the data product (node or reach), the simulation characteristics, and the particularities of aggregation. The validation results are presented here separately for different variables and data products.

## Node Results

Scaled node-level errors varied in behavior across the different simulations and variables. Height error RMSE was generally less than 1.3 for the realistic simulations--with most of this coming from bias--and less than 1.1 for the idealized simulations--which had negligible bias. Scaled width errors were more variable, with RMSE as high as 5x the model-predicted RMSE in the simple aggregation, 2x the predicted RMSE using composite aggregation, and 1.5x the predicted RMSE using water-fraction aggregation. The simulation parameters--idealized versus realistic--had only a minor effect on width error statistics.

```{r nodesdplot}
load("robjs/nodesdplot.RData")
load("robjs/nodesdplot_raw.RData")
nodesdplot
nodesdplot_raw
```


Normal quantile-quantile (QQ) plots (Fig. ####) compare the emprical scaled errors to the theoretical $N(0, 1)$ distribution (solid black line) for "water fraction" aggregtion and idealized simulation parameters. This distinguishes characteristics of the empirical error distribution including bias (vertical offset from 1:1 line), standard deviation (slope of points), and departures from normality, for instance heavy tails. Based on Fig. ####, the uncertainty model for height adequately characterized uncertainty across all simulation dates and for all parts of the distribution, with approximately standard normal scaled errors. In contrast, scaled width errors exhibit heavy-tailed distributions, especially at the upper tail, corresponding to errors that are much higher than would be expected if the errors were normally distributed. Despite this behavior in the tails, the middle of the scaled width error distribution was more well-behaved, with small but non-negligible bias depending on the simulation date. 

```{r qqplot_byday}
load("robjs/qqgg_nodes_frac_low.RData")
qqgg_nodes_frac_low
```

In aggregate across all simulation days, distributions of scaled node height and width errors exhibit broadly the same characteristics as the same distributions for individual simulation days, but differ in severity by depending on the simulation parameters (Fig. ####). Width errors were closest in distribution to standard normal for the "water fraction" aggregation method, while using the composite method resulted in a heavier right-skew, and the simple mehtod had both a severely large upper tail as well as higher variance throughout the distribution (Fig. ####a). Height error distributions were unaffected by aggregation method, but adding simulation noise resulted in significant bias and slightly larger variance (Fig. ####b. 

```{r qqplot_bycondition}
load("robjs/wseqq_compare.RData")
load("robjs/widthqq_compare_all3.RData")

widthqq_compare_all3
wseqq_compare

```


### Factors affecting node-level errors and uncertainty

Node-level height errors vary as a function of the number of pixels per node and the distance from the node to the ground-track of the satellite. These effects are not independent, since pixel size is larger closer to the ground track, resulting in fewer pixels for a given node area. A node with fewer pixels will have less ability to average out the independent noise per pixel, resulting in larger node-level errors. Pixel height errors also increase farther from the ground-track, as height becomes increasingly sensitive to changes in interferometric phase with increasing cross-track distance. These two phenomena--pixel size and height-phase sensitivity--result in a U-shaped relationship between estimated height uncertainty and cross-track distance (Fig. ####d). Both of these effects are well characterized in the height uncertainty model, resulting in scaled errors lying generally between -2 and 2, as expected given unit standard deviation.



```{r wse_scatter}
load("robjs/wseerr_gg1.RData")
wseerr_gg1
```


Width errors are also affected by pixel size, but are not otherwise affected by cross-track distance, resulting in the error and uncertainty profile shown in Fig. ####a-d. While the scaled width errors are not as well-behaved as the height errors and exceed 10 or 20 times the estimated $1\sigma$ uncertainty, they exhibit no discernable patterns when plotted against pixel count or cross-track distance, suggesting that these effects are well characterized by the uncertainty model. 


```{r width_scatter}
load("robjs/widtherr_gg1.RData")
widtherr_gg1
```


## Reach results 

Scaled reach-level errors for height and slope measurements (Fig. ####, top row) were apparently unbiased and fell generally between -2 and 2 in the idealized simulations, reflecting well-characterized uncertainties. As at the node level, corresponding width errors were less well-behaved, reaching as much as 10 times the $1\sigma$ uncertainty estimate. The magnitude of scaled width errors were larger on average than at the node scale, resulting from violations of underlying assumptions (zero bias, independence) when aggregating node to reach uncertainty.

In the "realistic" simulations, introducing simulation noise primarily affected height errors--biasing many simulations' reach errors--but also had non-negligible effects on slope and width errors (Fig. ####, bottom row). Slope errors were larger in magnitude in some reaches--notably reaches 6 and 8. Width errors varied more sporadically, increasing or decreasing depending on the simulation day and reach. 

```{r reach_scatter}
load("robjs/scatter_reach_final.RData")
scatter_reach_final
```



## Large width errors

Both the node-scale and reach-scale errors have right-skewed distributions, resulting from relatively infrequent sections of river having outsized influence on the aggregated errors. These errors result mainly from three distinct phenomena: complex channel geometry, spurious contiguity of water pixels, and narrow channels relative to pixel size. These are described in detail in the following text, and illustrated in Figure ####. 

1. Complex channel geometry that cannot be resolved, either due to pixel size or regularization in slant plane (Fig. ####a)
2. Sections of water that are contiguous in the PIXC slant plane but not in the validation slant plane, resulting in pixels for a given area of water being assigned to nodes in the PIXC that are not assigned to the node in the GDEM PIXC. (Fig. ####b)
3. Stretches of river that are narrow in comparison to the pixel size, with an apparent positive bias in water fraction estimates for water-edge pixels. (Fig. ####c). 

### Illustration of nodes with large errors

While pixel-level data do not have a 1:1 correspondence between simulated and validation data and thus cannot be compared directly, visual juxtaposition of pixel data illustrative of phenomena that lead to large width errors. These phenomena include:

1. Complex channel geometry that cannot be resolved, either due to pixel size or regularization in slant plane (Fig. ####a)
2. Sections of water that are contiguous in the PIXC slant plane but not in the validation slant plane, resulting in pixels for a given area of water being assigned to nodes in the PIXC that are not assigned to the node in the GDEM PIXC. (Fig. ####b)
3. Stretches of river that are narrow in comparison to the pixel size, with an apparent positive bias in water fraction estimates for water-edge pixels. (Fig. ####c)

All 3 of these phenomena are spatially correlated--nodes that experience them tend to occur in clusters. Thus the resulting errors at the reach scale are larger in magnitude, since node-to-reach uncertainty accumulation assumes that errors are spatially independent. 

```{r pix_maps}
load("robjs/pixmapgg.RData")
pixmapgg
```


